Querying Integrated Scientific Observation and Measurement Data

### 2010-05-07 docs/algdraft/DataIntegration/DataIntegration.tex
Refine the part for querying materialized database. The big structure of the paper is set. Need further refinement....

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Michael Shell}
\IEEEauthorblockA{School of Electrical and\\Computer Engineering\\
Georgia Institute of Technology\\
Atlanta, Georgia 30332--0250\\
Email: http://www.michaelshell.org/contact.html}
\and
\IEEEauthorblockN{Homer Simpson}
\IEEEauthorblockA{Twentieth Century Fox\\
Springfield, USA\\
Email: homer@thesimpsons.com}
\and
\IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
\IEEEauthorblockA{Starfleet Academy\\
San Francisco, California 96678-2391\\
Telephone: (800) 555--1212\\
Fax: (888) 555--1212}}

@@

\section{Introduction}
In this work, we study the query processing over scientific
observation and measurement data using OBOE
model\cite{DBLP:conf/er/BowersMS08}. OBOE model is a conceptual model
used to interpret observation and measurement data.

@@

%Challenge to query such data
\from{HP}{Queries to such data and challenges to answer such queries.}
When a user formulates a query, it is unrealistic for a him/her to know the 
underlying data structure of the data.  
But they all know the well recognized OM concepts as discussed above. 
So, naturally, when searching such scientific data, people are more
interested in finding datasets related to such OM terminologies as observation
measurement, characteristic, etc. 
Given the dataset in Table \ref{tb:dataset}. 
People may ask the following queries.

\begin{example}\label{eg:query}
Simple and summarization queries: 
\begin{itemize}
\item $Q_1$: Give me the data sets that contain species ``Picea rubens'' observations.
%$Q_2$: Give me the data sets that have measurements on ``area''  characteristics. 
\item $Q_2$: Give me the data sets that contain species ``Picea rubens''
  observations in ``California''. 
\item $Q_3$: Give me the data sets that contain at least five distinct
  ``Picea rubens'' observations.
\item $Q_4$: Give me the data sets that have trees with average ``height''
  than 20.0 in ``California''. 
% Shawn: an example on summarize one measure with respect to another.
% This could take two forms: one where the context forms the groups
% (like by year), and another where the two measurements are within the
% same context (like finding correlations).
\end{itemize}
\end{example}

@@

\from{HP}{Our effort in querying observational data: introduce OBOE, query}

@@

\subsection{Contribution and paper organization}
Contributions of this work:
\begin{itemize}
\item Formalize the queries over observation and measurement query.
\item Propose three methods to evaluate queries. 
\end{itemize}

This paper is organized as follows. 
Section \ref{sec:relatedwork} reviews works that are related to this
research. Section \ref{sec:dataquery} formalize the data model and the
queries that people are interested to ask.

@@

%What kinds of annotation information is needed.

We use $A$ to denote the annotation of one dataset. 
Internally, we keep the following information for the
annotation.

@@

\begin{itemize}
\item $A.OT$ = $\{(\underline{A_{id},ot_{id}}, et, isDistinct)\}$ to
  describe an observation type, which denotes on which
  entity type (object in the real world) the observation is made. 
Very often, more than one observation can be made on one entity. Thus,{\em
  isDistinct} is used to denote whether the same value of key
meaasurements of an obbservation types can uniquely identify one
observation or not. 
\item $A.MT$ = $\{(\underline{A_{id}, mt_{id}}, ot_{id},isKey,
  Cha, \cdots)\}$. Generally, $MT$ contains information about
  characteristic ($Cha$), Standard, Protocol and Precisio. We do not include them
  here to make the description clearer. Here, {\em isKey} is used to denote
  whether one measurement type is the key measurement for the
  observation type $OT_{id}$ or not. 
\item $A.CT$ = $\{(\underline{A_{id}, ot_{id}, cot_{id},
  Rel}, isIdentify)\}$ where $cot$ represents the context observation
  type. 
\item $A.Map$ = $\{(\underline{A_{id},mt_{id}, attr_i, Cond},
  Val\}$
\end{itemize}

@@

\begin{example}
$A.OT$ = \{($A_1, OT_1$, tree, false),($A_1, OT_2$, GeoSpot, true) \}
  denotes that annotation $A_1$ are for two observation types. One
  type $OT_1$ is on real world entity ``tree'' and the values of its key measurements
  do not uniquely identify a distinct observation. The other type
  $OT_2$ is for real world entity ``GeoSpot'' (i.e., geospatial
  location) and its key measurement types can identify it's unique
  observations. \from{HP}{Introduce key measurements before this.}

$A.MT$ =\{($A_1, MT_1, OT_1$, Species, true), 
($A_1, MT_2, OT_1$, SpecNo, true), 
($A_1, MT_3, OT_1$, Height, false),
($A_1, MT_4, OT_2$, Plot\_State, true), 
($A_1, MT_5, OT_2$, Plot\_Area, false)\} represents that $OT_1$
(for ``tree'') has three measurement types ($MT_1, MT_2$, $MT_3$) for characteristics
``Species'', ``SpecNo'', and ``Height''. The fist two measurement
  types together form the key measurement type for $OT_1$. 
$OT_2$ (for ``GeoSpot'') has one key measurement type $MT_4$ for
  characteristic ``Plot\_State'' and another measurement type $MT_5$
  for characteristic ``Plot\_Area''.

$A.CT$ = \{($A_1$,$OT_1$, $OT_2$, within, true)\}. It shows that
observation of a {\em tree} is within the context of a geo-spatial
location. And this context is used to identify the uniqueness of the
observation. E.g., a tree with species name ``piru'' and species no
$1$ in California
%$(34^{\circ}8'3N,118^{\circ}14'37W)$ (Los Angeles in California) 
is different from a tree with 
the same species name ``piru'' at Oregon. 
%location $36^{\circ}10'30N,115^{\circ}8'11W$ (Las Vegas in
%California).

@@

$A.Map$ = \{($A_1$, $MT_1$, ``code'', ``eq `piru''', ``Picea rubens''),
%($A_1$, $MT_1$, ``code'', ``eq `abba''', ``Abies balsamea''),
%($A_1$, $MT_1$, ``code'', ``eq `capo''', ``California poppy''),
($A_1$, $MT_2$,``no'', null,), 
($A_1$, $MT_3$, ``plt'', null,), 
($A_1$, $MT_4$,``area'', null,\}. 
The first mapping rules maps ``code''
attribute to the measurement type $MT_1$ (for ``Species''
characteristic) and change the value to ``Picea rubens'' if the code
is ``piru''. The meaning of the other mapping rules are very
obvious here, so we skip the details. 
\end{example}

\subsection{Observation and measurement (OM) query}
Above we defined the data model and the annotation, in this
sub-section, we define the queries that scientists are interested in
We denote the queries on the observational data as {\em Observation
and measurement (OM) query}.

In using the observational data, people are generally interested in
the following concepts: {\em observation, measurement,characteristic or standard}.

\begin{definition}[Basic OM query]\label{def:basic_omq}
A basic observation and measurement query is defined as $Q$::={\tt
  concept:cond}. 
Here, {\tt concept} is the main term in an observation model. 
{\tt cond} is in the form of $f([distinct]~attr)~op~value$ where $op$ is
basic operator from $\{=, \neq, >, <\}$ and $f$ is an
aggregation function from $\{sum, avg, count, min, max\}$. 
\end{definition}

\begin{definition}\label{def:qresult}
The result of a query $Q$ is a set of data objects (e.g., data tables) 
$\{d|d\in D \wedge d~satisfies~{\tt cond}\}$. 
For each of such result $d$, we use $d~s.t.~Q$ to denote that
data object $d$ satisfies the query. 
\end{definition}

Based on the different {\tt cond} definition, ``$d~satisfies~{\tt cond}$''
is translated into different formulas. 
E.g., if {\tt cond} is defined that the value of attribute ``area'' need to be
smaller than $3.0$, then, ``$d~satisfies~{\tt cond}$'' is translated to
``$dv<3.0|dv\in d[area]$''.

Definition \ref{def:basic_omq} can formulate queries on the
one specific observation and measurement. 
In real application, we need to consider the context relationship. 
So, we generalize the basic OM query to contextualized OM query as
follows.

\begin{definition}[Contextualized OM query]\label{def:context_omq}
$CQ$::=$Q_1~${\tt context}($Q_1.concept, Q_2.concept$)$~Q_2$. Here $Q_i$ is a basic query
in the form of {\tt concept:cond}. 
\end{definition}

@@

%\begin{example}
\begin{itemize}
\item $Observation: \langle Species='Picea~rubens'\rangle$
\item $Observation_1: \langle Species='Picea~rubens'\rangle$\\
  $IN(Observation_1, Observation_2$) \\
$Observation_2: \langle Meas_{unknown}=`California'\rangle$
\item $Observation: \langle Species='Picea~rubens' $\\
$\wedge
  count(distinct~Species)\geq 5\rangle$
\item $??$
\end{itemize}

@@

In this section, we discuss several methods that we can use to
answer OM queries over scientific observational data with
annotations.

%\subsection{Naive method}\label{sec:naive}

A very direct way search such data is to extract all the content in
the data cells and index them, this way, the system may help answer
very simple queries with the help of some thesaurus. 
For example, for $Q_1$ in Example \ref{eg:query}, as long as we can
translate the abbreviation ``piru'' in the dataset to ``Picea rubens''
based on some predefined thesaurus. We can find the results. 
Even for this kind of simple query, however, if we want to find
observation of ``California poppy'', an approximate match may return
the dataset in Table \ref{tb:dataset} as a result because it contains
both {\em California} and {\em Poppy}.
If such compound word search can be alleviated using some existing
techniques \cite{***}, %citation on compound word querying. 
it is almost impossible to directly use such IR-style system to answer a query
in the form of $Q_3$ and $Q_4$. 
I.e., answer queries with context.

This naive method cannot answer the desired query. It is because it
fails to catch the semantics in the data. 
In what follows, we would use the annotation during our search to get
more semantic support.

As we analyzed in Section \ref{sec:relatedwork}, there are two extrems
in querying integrated data. 
One is to rewrite the original query to a series of new queries over
the data; the other is to materialize the data to a consistent data
model and then answer the query over the materialized database. 
We what follows we work from these two directions to leverage the
semantic information in answering such queries.

@@

[
{\tt 
SELECT DISTINCT d.record\_id\\
FROM d\\
WHERE $f(d_{a_1} \cdots d_{a_m}) = true$\\
$[$GROUP BY$]$\\
$[$HAVING$]$
}
]

@@

From the above description, we can tell that the computation cost is to search the different data
tables. This way, for each candidate data table, we need to send an
SQL query to the server for evaluation purpose.
If the number of candidate data tables are small. This method should
work well. However, when we have a lot of candidate data tables, the
efficiency may be affected. 
In what follows, we introduce another strategy, which logically merge
the data content in different tables and perform queries ober the
materialized database.

\subsection{Querying over materialized database}\label{sec:matdb}

@@

The basic idea is to materialize the data into some centralized
concept tables with instances of entities, observations, measurements
and the contexts between observations. 
We call such concept tables core OBOE tables, and denote them as
$OBOE.*$. 
In partibular, we have the following concept instance tables. 
\begin{itemize}
\item $OBOE.EI$ = $\{(\underline{ei_{id}}, et_{id})\}$ for all the entity instances;
\item $OBOE.OI$ = $\{(\underline{oi_{id}}, ot_{id}, ei_{id})\}$ keeps
  all the observation instances;
\item $OBOE.MI$ = $\{(\underline{mi_{id}}, oi_{id}, mt_{id},
  mVal)\}$ for all the measurement instances;
\item $OBOE.CI =\{(\underline{oi_{id},coi_{id},ct_{id}})\}$ for
  all the context instances;
\end{itemize}

In what follows, we describe how to materialize the data into such
concept instance tables in Section \ref{sec:materialilzedb} when
considering all the distinct and context constratins. 
Then, in Section \ref{sec:materialilzedb}, we describe our strategy to
perform queries over materialized database.

@@

\subsubsection{Querying materialized database}\label{sec:materialilzedb}

\begin{example} \label{eg:materialize_db_q1}
Example \ref{eg:query_reqerite_q1} shows how to deal with $Q_1$ using
the query rewriting method. 
Using the  materialized database, the first step is the same.
After we know the data file, then we search the OBOE.MI with the
condition that {\em mVal=``Picea rubens''}.
\end{example}

{\tt 
SELECT distinct $AD.d_{id}$\\
FROM MI, MT, AD\\ 
WHERE $MT.Cha='Species'$ \\
AND $MI.mVal = `Picea rubens'$\\
AND $MI.mt_{id}=MT.mt_{id}$ AND $MT.A_{id}=AD.A_{id}$; %\#Join the MI and MT tables
}

In this example, the computation cost is the selection cost of the
measurement instance table. 
It should be faster than the query rewriting.

If the query has some restriction on the aggregate result, we can
easily add the GROUP BY and HAVING clause to deal with it. 
We take $Q_3$ as an example, which asks for data sets that contain at
least five distinct ``Picea rubens'' observations.
The SQL query can be written as: \\
{\tt 
SELECT distinct $d_{id}$\\
FROM MI, MT\\ 
WHERE $MT.Cha='Species'$ \\
AND $MI.mVal = `Picea rubens'$\\
AND $MI.mt_{id}=MT.mt_{id}$\\
GROUP BY $d_{id},oi_{id}$\\
HAVING COUNT(*)$>$5; %\#Join the MI and MT tables
}

Using the query rewriting, 
First, we find the data tables that contain the needed concepts. 
In each table, we need to figure out the distinct observations. 
From $m_i$ $obstype_i$, all the other measurements that are of
$obstype_i$. 
Group by the key measurements of $obstype_i$ if distinct yes.
If this observation type does not have any context, then the key
measurements are all the measurements directly defined under this
observatioin type. Otherwise, i.e., this observation type has context,
then the key measurements are all the measurements in the context
chain. 
%\end{example}

When we are given contextualized query, e.g., $Q_2$. 
Give me the datasets that contain species ``Picea rubens'' observations in ``California''.

Species = ``Picea rubens'' and observation has context {\em in}
another observation which has measurement of value ``California''.

@@

%select mtype, olabel from measurement\_type  \\
%where chracteristic = 'Species';  \\

{\tt 
SELECT record\_id\\
FROM CI as ci, MI as mi, MI as cmi,MT as mt, MT as cmt, \\
WHERE $mi.oi_{id}=ci.oi_{id}$ AND $cmi.oi_{id}=ci.coi_{id}$ \\
AND $ci.coi_{id} = coi.oi_{id}$ AND oi.otype = 'olabel' \\
AND mi.mvalue=`Picea rubens' AND cmi.mvalue='California'\\
AND MT.mt_{id}=mi.mt_{id} AND MT.mLabel = 'Species'\\
AND CMT.mt_{id}=cmi.mt_{id} AND cmt.mLabel=`State'\\
AND coi.otype IN (\\
\verb|    |SELECT context\_olabel\\
\verb|    |FROM context\_relationship\\
\verb|    |WHERE olabel = 'olabel');
}
The sub-query finds the set containing all the observation type
labels in the context chain.

Use query rewriting, what I can do? 
It's the same till to getting context\_olabel\_set.

>>>>

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Author1}
\IEEEauthorblockA{School of Electrical and\\Computer Engineering\\
Georgia Institute of Technology\\
Atlanta, Georgia 30332--0250\\
Email: http://www.michaelshell.org/contact.html}
\and
\IEEEauthorblockN{Author2}
\IEEEauthorblockA{Twentieth Century Fox\\
Springfield, USA\\
Email: homer@thesimpsons.com}
\and
\IEEEauthorblockN{Author3}
\IEEEauthorblockA{Starfleet Academy\\
San Francisco, California 96678-2391\\
Telephone: (800) 555--1212\\
Fax: (888) 555--1212}}

@@

\section{Introduction}\label{sec:intro}
In this work, we study the query processing over scientific
observation and measurement data using OBOE
model\cite{DBLP:conf/er/BowersMS08}. OBOE model is a conceptual model
used to interpret observation and measurement data.

@@

%Queries to such data and challenges to answer such queries.
The internal data structure of such scientific data repository is
generally unknown to a user who wants to pose a query to such scientific data
repository and get some useful information. 
So, it is unrealistic for a him/her to formulate a query based on the 
underlying data structure of the data.  
The easiest query they can pose to such data repository are keyword-query. 
Such keyword style query works well for the dataset metadata
(e.g., contributor information, dataset description, etc) using the
current Information Retrieval (IR) techniques. However, it
is far from enough to be used to search the data content because the
data content need to be interpreted with semantics. 
It is important in the scientific domain to have typical
kinds of queries. One fact we can use is that the domain
scientists know well the widely used OM concepts as discussed above. 
So, naturally, when searching such scientific data, they can provide
the concept information to restrain their queries and to find data sets
related to such OM terminologies. 
Given the dataset in Table \ref{tb:dataset}. 
People may ask the following observation and measurement (OM) queries.

\begin{example}\label{eg:query}
Simple and summarization queries: 
\begin{itemize}
\item $Q_1$: Give me the data sets that contain species ``Picea rubens'' observations.
%$Q_2$: Give me the data sets that have measurements on ``area''  characteristics. 
\item $Q_2$: Give me the data sets that contain species ``Picea rubens''
  observations in ``California''. 
\item $Q_3$: Give me the data sets that contain at least five distinct
  ``Picea rubens'' observations.
\item $Q_4$: Give me the data sets that have trees with average ``height''
  than 20.0 in ``California''. 
% Shawn: an example on summarize one measure with respect to another.
% This could take two forms: one where the context forms the groups
% (like by year), and another where the two measurements are within the
% same context (like finding correlations).
\end{itemize}
\end{example}
\from{HP}{Challenges to answer such queries: uniqueness}

@@

\from{HP}{Our effort in querying observational data: introduce OBOE}

@@

\subsection{Contribution and paper organization}
Contributions of this work:
\begin{itemize}
\item We formulate a canonical set of queries over observation and
  measurement scientific data repository. Such formal queries can
  formalize most OM concepts related searches. 
\item We propose three methods to evaluate such OM
  queries. \from{HP}{Elaborate the three methods.}
\end{itemize}

This paper is organized as follows. 
Section \ref{sec:relatedwork} reviews works that are related to this
research. Section \ref{sec:dataquery} formalizes the data model and the
queries that people are interested to ask.

@@

%What kinds of annotation information is needed. 
We use $A$ to denote the annotation of one dataset. 
Internally, we keep the following information for the
annotation.

@@

\begin{itemize}
\item $OT$ = $\{(\underline{A_{id},ot_{id}}, et, isDistinct)\}$ to
  describe an observation type, which denotes on which
  entity type (object in the real world) the observation is made. 
Very often, more than one observation can be made on one entity. Thus,{\em
  isDistinct} is used to denote whether the same value of key
meaasurements of an obbservation types can uniquely identify one
observation or not. 
\item $MT$ = $\{(\underline{A_{id}, mt_{id}}, ot_{id},isKey,
  Cha, \cdots)\}$. Generally, $MT$ contains information about
  characteristic ($Cha$), Standard, Protocol and Precision. We do not include them
  here to make the description clearer. Here, {\em isKey} is used to denote
  whether one measurement type is the key measurement for the
  observation type $OT_{id}$ or not. 
\item $CT$ = $\{(\underline{A_{id}, ct_{id}}, ot_{id}, cot_{id}, Rel,
  isIdentify)\}$ where $cot$ represents the context observation
  type. 
\item $Map$ = $\{(\underline{A_{id},mt_{id}, attr_i, MapCond},
  Val\}$
\end{itemize}

@@

\begin{example}
$OT$ = \{($A_1, OT_1$, tree, false),($A_1, OT_2$, GeoSpot, true) \}
  denotes that annotation $A_1$ are for two observation types. One
  type $OT_1$ is on real world entity ``tree'' and the values of its key measurements
  do not uniquely identify a distinct observation. The other type
  $OT_2$ is for real world entity ``GeoSpot'' (i.e., geospatial
  location) and its key measurement types can identify it's unique
  observations. \from{HP}{Introduce key measurements before this.}

$MT$ =\{($A_1, MT_1, OT_1$, Species, true), 
($A_1, MT_2, OT_1$, SpecNo, true), 
($A_1, MT_3, OT_1$, Height, false),
($A_1, MT_4, OT_2$, Plot\_State, true), 
($A_1, MT_5, OT_2$, Plot\_Area, false)\} represents that $OT_1$
(for ``tree'') has three measurement types ($MT_1, MT_2$, $MT_3$) for characteristics
``Species'', ``SpecNo'', and ``Height''. The fist two measurement
  types together form the key measurement type for $OT_1$. 
$OT_2$ (for ``GeoSpot'') has one key measurement type $MT_4$ for
  characteristic ``Plot\_State'' and another measurement type $MT_5$
  for characteristic ``Plot\_Area''.

$CT$ = \{($A_1$,$OT_1$, $OT_2$, within, true)\}. It shows that
observation of a {\em tree} is within the context of a geo-spatial
location. And this context is used to identify the uniqueness of the
observation. E.g., a tree with species name ``piru'' and species no
$1$ in California
%$(34^{\circ}8'3N,118^{\circ}14'37W)$ (Los Angeles in California) 
is different from a tree with 
the same species name ``piru'' at Oregon. 
%location $36^{\circ}10'30N,115^{\circ}8'11W$ (Las Vegas in
%California).

@@

$Map$ = \{($A_1$, $MT_1$, ``code'', ``eq `piru''', ``Picea rubens''),
%($A_1$, $MT_1$, ``code'', ``eq `abba''', ``Abies balsamea''),
%($A_1$, $MT_1$, ``code'', ``eq `capo''', ``California poppy''),
($A_1$, $MT_2$,``no'', null,), 
($A_1$, $MT_3$, ``plt'', null,), 
($A_1$, $MT_4$,``area'', null,\}. 
The first mapping rules maps ``code''
attribute to the measurement type $MT_1$ (for ``Species''
characteristic) and change the value to ``Picea rubens'' if the code
is ``piru''. The meaning of the other mapping rules are very
obvious here, so we skip the details. 
\end{example}

\subsection{Observation and measurement (OM) query}
Above we defined the data model and the annotation. In this
sub-section, we formalize the queries that scientists are interested
in. We denote the queries on the observational data as {\em Observation
and measurement (OM) query}.

As discussed in Section \ref{sec:intro}, it is generally impossible for a
scientist to formulate an OM query using the internal data structure
of a dataset because such data structures differ from contributors to
contributors. More realistically, they can pose queries
(Example \ref{eg:query}) related to key OM concepts, e.g., {\em
  observation, measurement,characteristic, and standard}, and these
OM concepts satisfy given keyword and context conditions.

\begin{definition}[Basic OM query]\label{def:basic_omq}
A basic observation and measurement query is defined as 
\[Q::={\tt OMconcept:}\langle{\tt cond}\rangle}\]. 
Here, 
\begin{itemize}
\item {\tt OMconcept} refers to one of the major OM terminologies. 
\item 
$\langle${\tt cond}$\rangle$ is in the form of 
{\tt f([distinct entity.]measurement) op value} 
where {\tt op} is a basic operator from $\{=, \neq, >, <\}$ 
and {\tt f} is an
aggregation function from $\{sum, avg, count, min, max\}$. 
{\tt distinct} and {\tt entity} are optional. 
\end{itemize}
\from{HP}{The symbol fonts are a little bit messy...need to
  distinguish between the keywords and variables.}
\end{definition}

\begin{definition}[Result of OM query]\label{def:qresult}
The result of a query $Q$ is a set of data objects (e.g., data tables) 
\{$d|d\in D \wedge d~satisfies~\langle ${\tt cond} $\rangle$\}. 
For each of such result $d$, we use $d~s.t.~Q$ to denote that
an data object $d$ satisfies the query. 
\end{definition}

Based on the different $\langle${\tt cond}$\rangle$ definitions,
``$d~satisfies~\langle${\tt cond}$\rangle$''
is translated differently. 
E.g., if $\langle${\tt cond}$\rangle$ is defined that the value of ``{\em area}'' need to be
smaller than $3.0$, then, ``$d~satisfies~{\tt cond}$'' is translated to
``$dv<3.0|dv\in d[area]$''.

Definition \ref{def:basic_omq} can formulate queries on
one specific observation and measurement. 
In real application, we need to consider the context relationship. 
\from{HP}{E.g., etc.}
So, we generalize the basic OM query to contextualized OM query as
follows.

\begin{definition}[Contextualized OM query]\label{def:context_omq} A
  contextualized OM query is defined as:
\[CQ ::=Q_i \wedge {\tt context}(Q_i.{\tt OMconcept}, Q_j.{\tt
  OMconcept})\wedge Q_j\]
Here, $Q_i, Q_j$ are basic queries.
\end{definition}

@@

%\begin{example}
\begin{itemize}
\item $Observation: \langle Species='Picea~rubens'\rangle$
\item $Observation_1: \langle Species='Picea~rubens'\rangle$\\
  $\wedge IN(Observation_1, Observation_2$) \\
$\wedge Observation_2: \langle Meas_{unknown}=`California'\rangle$
\item $Observation: \langle Species='Picea~rubens'$\\
$\wedge count(distinct~Species)\geq 5\rangle$
\item $Observation_1: \langle avg(distinct~tree.height)\geq 20.0 \rangle$ 
$\wedge IN(Observation_1, Observation_2)$\\
$\wedge Observation_2: \langle State='California'\rangle$\\
%Give me the data sets that have trees with average ``height''
%  than 20.0 in ``California''
\end{itemize}

@@

In this section, we propose several methods that can be used to
answer OM queries over scientific observational data with
annotations.

%\subsection{Naive method}\label{sec:naive}
A very direct way to search such data is to extract all the content in
the data cells and index them. This way the system may help answer
very simple queries with the help of some thesaurus. 
For example, for $Q_1$ in Example \ref{eg:query}, as long as we can
translate the abbreviation ``piru'' in the dataset to ``Picea rubens''
based on some predefined thesaurus, we can find the results. 
However, even for this kind of simple query, if we want to find
observation of ``California poppy'', an approximate match may return
the dataset in Table \ref{tb:dataset} as a result because it contains
both {\em California} and {\em Poppy}.
If the search problem caused by the compound words can be alleviated using some existing
techniques \cite{***}, %citation on compound word querying. 
it is still almost impossible to directly use such IR-style system to answer a query
in the form of $Q_3$ and $Q_4$. 
This naive method cannot be directly used answer the desired queries. It is because it
fails to catch the semantics of the context and uniqueness of observations in the data. 
In what follows, we utilize the annotation information to get
more semantic support, and provide new strategies to answer the
queries.

As we analyzed in Section \ref{sec:relatedwork}, there are two extrems
in querying integrated data. 
One is to rewrite the original query to a series of new queries over
the data; the other is to materialize the data to a consistent data
model and then answer the query over the materialized database. 
In what follows we work from these two directions to leverage the
semantic information in answering such queries.

@@

\vspace{0.1in}
{\tt 
\begin{tabular}{l} 
SELECT DISTINCT $d_{id}$\\
FROM $d$\\
WHERE $f(d_{a_1} \cdots d_{a_m}) = true$\\
$[$GROUP BY {\em distinct requirement}$]$\\
$[$HAVING {\em aggregation condition}$]$;
\end{tabular}
}
\vspace{0.1in}

@@

\from{HP}{Refine the process of $Q_2$.\\
For the first step, we find the data tables that contain the needed concepts. 
In each table, we need to figure out the distinct observations. 
From $m_i$ $obstype_i$, all the other measurements that are of
$obstype_i$. 
Group by the key measurements of $obstype_i$ if distinct yes.
If this observation type does not have any context, then the key
measurements are all the measurements directly defined under this
observatioin type. Otherwise, i.e., this observation type has context,
then the key measurements are all the measurements in the context
chain. }

%Analysis of this method.

{\bf Execution cost analysis:} 
From the above description, we can tell that the computation cost is to search the different data
tables. This way, for each candidate data table, we need to send an
SQL query to the server for evaluation purpose.
If the number of candidate data tables are small. This method should
work well. However, when we have a lot of candidate data tables, the
efficiency may be affected. 
In what follows, we introduce another strategy, which logically merge
the data content in different tables and perform queries ober the
materialized database.

\subsection{Querying materialized database}\label{sec:matdb}

@@

The basic idea is to materialize the data into some centralized
concept tables with instances of entities, observations, measurements
and the contexts between observations. 
We call such concept tables core OBOE tables, and denote them as
$OBOE.*$. 
In partibular, we have the following concept instance tables. 
\begin{itemize}
%\item $EI$ = $\{(\underline{ei_{id}}, et_{id})\}$ for all the entity instances;
\item $OI$ = $\{(\underline{oi_{id}}, ot_{id}, ei_{id})\}$ keeps
  all the observation instances where $ot_{id}$ is used to link to the
  annotation information in $OT$; 
\item $MI$ = $\{(\underline{mi_{id}}, oi_{id}, mt_{id},
  mVal)\}$ for all the measurement instances where $mt_{id}$ is used to link to the
  annotation information in $MT$; 
\item $CI =\{(\underline{oi_{id},coi_{id},ct_{id}})\}$ for
  all the context instances where $ct_{id}$ is used to link to the
  annotation information in $CT$. 
\end{itemize}

In what follows, we describe how to materialize the data into such
concept instance tables in Section \ref{sec:materialilzedb} when
considering all the distinct and context constratins. 
Then, in Section \ref{sec:query_materialilzedb}, we describe our strategy to
perform queries over materialized database.

@@

\subsubsection{Querying materialized database}\label{sec:query_materialilzedb}

Once the data are materialized using the annotation constraints, we
can perform the query over them. In what follows, we describe how to
use the materialized data to answer the query.

When the given query is a basic OM query, i.e., there is not context
in the query. 
Then, we can form a template query as the following;

\begin{figure}[htb]
{\tt
\begin{tabular}{l}
SELECT $[$DISTINCT$]$ $AD.d_{id}$\\
FROM $AD, MI~as~mi, MT~as~mt[$, {\em other tables}$]$\\
$[$WHERE    {\em table join condition, \\
$\qquad\qquad$ AND selection condition}$]$,\\
$[$GROUP BY {\em distinct requirement} $]$\\
$[$HAVING   {\em aggregation condition} $]$;
\end{tabular}
}
\caption{SQL template to answer OM using materialized DB}
\label{fig:sql_matdb_basic_omq}
\end{figure}

In particular,

\vspace{0.1in}
\begin{tabular}{l}
$table~join~condition=$\\
%$\qquad (mi.oi_{id}=ci.oi_{id})~AND~(cmi.oi_{id}=ci.coi_{id})~AND$\\
$\qquad (mi.mt_{id}=mt.mt_{id})~AND~(mt.A_{id}=AD.A_{id})$
\end{tabular}
\vspace{0.1in}

For the very basic query (e.g., $Q_1$ in Example \ref{eg:query}), the
SQL generally need only the first three clauses. 
If the query has some restriction on the aggregated result, the {\tt
  GROUP BY} and {\tt HAVING} clauses to deal with it.

We use $Q_1$ and $Q_3$ to illustrate the query on the materialized
database. 
can be answered using the above
template. 
Example \ref{eg:query_reqerite_q1} shows how to deal with $Q_1$ using
the query rewriting method.

\begin{example} \label{eg:materialize_db_q1}
Using the  materialized database,  we just need to search the MI table
with the condition that {\em mVal=``Picea rubens''}.

\vspace{0.1in}
{\tt 
\begin{tabular}{l}
SELECT DISTINCT $AD.d_{id}$\\
FROM $AD, MI~as~mi, MT~as~mt$\\ 
WHERE $mt.Cha='Species'$ \\
$\qquad\quad$ AND $mi.mVal = `Picea~rubens'$\\
$\qquad\quad$ AND $mi.mt_{id}=mt.mt_{id}$\\
$\qquad\quad$ AND $mi.A_{id}=AD.A_{id}$; 
%\#Join the MI and MT tables
\end{tabular}
}
\vspace{0.1in}

For $Q_3$, which asks for data sets that contain at
least five distinct ``Picea rubens'' observations.
The SQL query can be written as:

\vspace{0.1in}
{\tt 
\begin{tabular}{l}
SELECT distinct $AD.d_{id}$\\
FROM $AD, MI~as~mi, MT~as~mt$\\
WHERE $mt.Cha='Species'$ \\
$\qquad\quad$ AND $mi.mVal = `Picea rubens'$\\
$\qquad\quad$ AND $mi.mt_{id}=mt.mt_{id}$\\
$\qquad\quad$ AND $mt.A_{id}=AD.A_{id}$\\
GROUP BY $d_{id},oi_{id}$\\
HAVING COUNT(*)$>$5; 
}
\end{tabular}
}
\vspace{0.1in}
\end{example}

For this kind of query, the computation cost is the selection cost of the
measurement instance (MI) table. 
Compared with the query rewriting strategy, which need to pose the
query over multiple data tables, this query should answer the queries
more efficiently.

Figure \ref{fig:sql_matdb_basic_omq} shows the SQL template to process
the basic OM query without context. In case the OM query has context,
we need to add the context into consideration.

\begin{enumerate}
\item The first step is similar to that in the basic OM query
  processing. Besides this, we need to figure out what are the concept
  obserbationi types of the needed observations.  
\item \label{step:singlesql} Form an SQL for the observation and every of its context
  observations. 
\item Intersect all the SQLs from Step \ref{step:singlesql}). 
\end{enumerate}

\begin{figure}[htb]
{\tt
\begin{tabular}{l}
SELECT $[$DISTINCT$]$ $AD.d_{id}$\\
FROM $AD, CI~as~ci,MI~as~mi, MT~as~mt$\\
$\qquad MI~as~cmi, MT~as~cmt[,$ {\em other tables}$]$\\
$[$WHERE    {\em table join condition, \\
$\qquad\qquad$ AND selection condition}$]$,\\
$[$GROUP BY {\em distinct requirement} $]$\\
$[$HAVING   {\em aggregation condition} $]$;
\end{tabular}
}
\caption{SQL template to answer contextualized OM using materialized DB}
\label{fig:sql_matdb_context_omq}
\end{figure}

In this case, the table join condition is much more complex since we
need to use the context instance table $ci$ to connect the
observation and context observation table. 
In particular,

\vspace{0.1in}
{\tt 
\begin{tabular}{l}
$table~join~condition=$\\
$\qquad (mi.oi_{id}=ci.oi_{id})$ AND $(cmi.oi_{id}=ci.coi_{id})$ AND\\
$\qquad (mi.mt_{id}=mt.mt_{id})$ AND $(cmi.mt_{id}=cmt.mt_{id})$ AND\\
$\qquad (mt.A_{id}=AD.A_{id})$
\end{tabular}
}
\vspace{0.1in}

{\bf Execution cost analysis:} From the join condition, we can see that the expensive cost comes from
the join operation between the measurement instance, context
measurement instance, and context instance table. 
As we analyzed in Section \ref{sec:materialilzedb}, the two tables with
growing space are measurement instance table and context instance
table.

{\bf Example:} When we are given contextualized query, e.g., $Q_2$, which asks for
the datasets that contain species ``Picea rubens'' observations in
``California''.

In the first step, we get the requirement is on Characteristic
``Species'' with value ($mVal$) ``Picea rubens'' and observation has context {\em in}
another observation for ``State'' of value ``California''.

@@

%select mtype, olabel from measurement\_type  \\
%where chracteristic = 'Species';  \\
In the first step, we need to find the context type id $tmp\_ct\_id$
that satisfy the context relationship.
